\part{The Chapters and Sections}
\markboth{Chapters and Sections}{Chapters and Sections}
\chapter*{Chapter~\ref{C:intro} \\Introduction to Writing in Mathematics}

\section*{Main Objectives}
After this chapter, students should:
\begin{itemize}
\item Understand the definition of a statement in mathematics.
\item Understand how mathematicians use conditional statements.
\item Understand the closure properties of the standard number systems.
\item Have begun the process of constructing and writing simple, direct proofs.
\item Understand and be able to use some of the conventions used in writing mathematics.

\end{itemize}
%\begin{itemize}
%\item Get the students started in writing mathematics.
%\item Introduce students to the definition of a statement in mathematics.
%\item Begin work on understanding conditional statements.
%\item Start the process of constructing and writing simple, direct proofs.
%\end{itemize}

It must be emphasized that the goal is to get students started with writing proofs and to understand how conditional statements are used in mathematics.  It is not expected that they will perfect their writing skills and their understanding of conditional statements at this time.  However, it is extremely important that students have a thorough understanding of conditional statements before starting Chapter~\ref{C:proofs}.  More work on conditional statements will occur in Chapter~\ref{C:logic}.
\hbreak

\input{manual/sections/section11}
\input{manual/sections/section12}













\chapter*{Chapter~\ref{C:logic} \\Logical Reasoning}

\section*{Main Objectives}
After this chapter, students should:
\begin{itemize}
\item Have a thorough understanding of the logical operations of negation, conjunction, and disjunction.
\item Have a thorough understanding of conditional statements in mathematics.
\item Have a thorough understanding of the use of quantifiers in mathematics.
\item Know some basic logical equivalencies, especially those used later in the text for methods of proof.
\item Be able to establish logical equivalencies.
\item Be able to write negations of statements with quantifiers.
\end{itemize}
%\begin{itemize}
%\item Provide students with a thorough understanding of the logical operations of negation, conjunction, and disjunction.
%\item Provide students with a thorough understanding of conditional statements in mathematics.
%\item Provide students with a thorough understanding of the use of quantifiers in mathematics.
%\item Provide students with an understanding of logical equivalencies, especially those used later in the text for methods of proof.
%\item Provide students with the ability to establish logical equivalencies.
%\end{itemize}
\hbreak
\input{manual/sections/section21}
\input{manual/sections/section22}
\input{manual/sections/section23}
\input{manual/sections/section24}
%\input{xsection23}
%\input{xsection21}
%\input{xsection24}


\chapter*{Chapter~\ref{C:proofs} \\Constructing and Writing Proofs in Mathematics}

\section*{Main Objectives}
\begin{itemize}
\item Provide students with a thorough understanding of the logical foundation of the following methods of proof:
\begin{multicols}{2}
\begin{list}{}
\item Direct proof;
\item Proof using the contrapositive;
\item Proof by contradiction;
\item Proof using cases.
\end{list}
\end{multicols}
\item Develop students' abilities to construct and write proofs using these and other proof methods.
\item Introduce students to the relations of ``divides'' and congruence.
\item Develop students' abilities to construct and write proofs dealing with divisors and congruence.
\item Develop students' understanding of the Division Algorithm and their ability to use the Division Algorithm in proofs using cases.
\item Develop students' understanding of the distinction between a constructive proof and a non-constructive proof.
\end{itemize}
\hbreak
\input{manual/sections/section31}
\endinput

\input{manual/sections/xsection32}
\input{xsection33}
\input{xsection34}
\input{xsection35}


\chapter*{Chapter~\ref{C:settheory} \\Set Theory}

\section*{Main Objectives}
\begin{itemize}
\item Provide students with a thorough understanding of the common definitions and notations used in set theory.
\item  Introduce students to the concepts of the power set of a set and the cardinality of a finite set.
\item Provide students with a thorough understanding of the common definitions and notations used to define the basic set operations.
\item  Insure that students understand Venn diagrams and are proficient in using Venn diagrams to explore possible set relationships.
\item  Develop students' ability to use the so-called ``Choose Method'' to prove set relationships and other results in mathematics.
\item Develop students' understanding of the ``Algebra of Sets.''  Develop students' ability to use the algebra of sets to prove set relationships.
\item Develop students' understanding of the Cartesian product of two sets and develop their ability to prove results about Cartesian products of sets.
\item Develop students' ability to work with indexed families of sets and operations on indexed families of sets.
\end{itemize}
\hbreak

\input{xsection41}
\input{xsection42}
\input{xsection43}
\input{xsection44}
\input{xsection45}

\chapter*{Chapter~\ref{C:induction} \\Mathematical Induction}

\section*{Main Objectives}
\begin{itemize}
\item Provide students with an understanding of the process of mathematical induction.
\item Develop the ability to construct and write proofs using mathematical induction.
\item Provide an understanding of the various forms of mathematical induction.
\item Provide an understanding of definition by recursion.
\item Develop the ability to construct and write proofs about sequences that are defined recursively.
\item Gain an understanding of the Fibonacci numbers.
\end{itemize}
\hbreak

\input{xsection51}
\input{xsection52}
\input{xsection53}

\chapter*{Chapter~\ref{C:functions} \\Functions}

\section*{Main Objectives}
\begin{itemize}
\item To understand the concept of a function including the related notation and terminology, such as domain, codomain, and range.
\item To be able to work with functions in a variety of settings and be able to recognize certain mathematical processes as functions.
\item To understand the definitions of some of the special types of functions such as injections, surjections, and bijections.
\item To be able to prove or disprove that specific functions are injections or surjections.
\item To be able to work with the concepts of injections and surjections in proofs.
\item To understand the concept of the composition of two functions and to be able to prove results dealing with the composition of functions.
\item To understand the ordered pair reprentation of a function.
\item To understand the definition of the inverse of a function and to be able to determine when the inverse of a function is itself a function.
\item To be able to write proofs of results dealing with the inverse of a function.
\end{itemize}
\hbreak

\input{xsection61}
\input{xsection62}
\input{xsection63}
\input{xsection64}
\input{xsection65}
\input{xsection91}


\chapter*{Chapter~\ref{C:equivrelations} \\Equivalence Relations}

\section*{Main Objectives}
\begin{itemize}
\item Understand the definition of a relation and be able to use the related notation and terminology, such as the domain and range of a relation.
\item Be able to explain why a function is a relation to to be able to give examples of relations that are not functions.
%\item Understand some of the standard mathematical relations.
\item Understand the definition of the inverse of a relation and to be able to prove results about relations and the inverse of a relation.
\item Understand the properties of reflexive, symmetric, or transitive relations and to be able to determine which of these properties a certain relation satisfies.
\item Understand the definition of an equivalence relation and know some standard equivalence relations including congruence modulo $n$.
\item Be able to determine whether or not a given relation is an equivalence relation.
\item Understand what an equivalence relation is and know the basic properties of equivalence classes.
\item Be able to prove some of the basic properties of equivalence classes and be able to explain why these properties mean that the equivalence classes form a partition of the underlying set.
\item Understand the definitions of addition and multiplication for the equivalence classes that form the integers modulo $n$.
\item To be able to explain the issue of ``well-defined'' when making a definition of addition or multiplication of equivalence classes.
\item Be able to use modular arithmetic to prove properties of the integers, including certain divisibility tests.
\end{itemize}
\hbreak

\input{xsection71}
\input{xsection72}
\input{xsection73}
\input{xsection74}

\chapter*{Chapter~\ref{C:numbertheory} \\Topics in Number Theory}

\section*{Main Objectives}
\begin{itemize}
\item Understand the properties of the greatest common divisor of two integers.
\item Be able to use the Euclidean Algorithm to compute the greatest common divisor of two integers and to write the greatest common divisor as a linear combination of the two integers.
\item Be able to use the properties of the greatest common divisor of two integers to prove results in number theory.
\item Understand the proof of the Fundamental Theorem of Arithmetic.
\item Solve linear Diophantine equations in two variables.
\item Understand the proof of the theorem which states the facts about solutions of linear Diophantine equations in two variables.
\end{itemize}
\hbreak

\input{xsection81}
\input{xsection82}
\input{xsection83}


\chapter*{Chapter~\ref{C:topicsinsets} \\Topics in Set Theory}

\section*{Main Objectives}
\begin{itemize}
\item Know the definition of a finite set and some of the basic properties of finite sets.
\item Understand the Pigeonhole Principle and be able to use it in appropriate settings.
\item Understand and be able to prove basic facts about infinite sets and countably infinite sets.
\item Be able to explain why $\mathbb{Z}$ and $\mathbb{Q}$ are countably infinite.
\item Be able to use Cantor's diagonal argument to explain why the open interval 
$\left( 0, 1 \right)$ is uncountable.
\item Know and be able to prove Cantor's Theorem that a set and its power set do not have the same cardinality.

\end{itemize}
\hbreak

\input{xsection92}
\input{xsection93}
\input{xsection94}

\endinput

