\subsection*{Proving that Something Does Not Exist}
In mathematics, we sometimes need to prove that something does not exist or that something is not possible.  Instead of trying to construct a direct proof, it is sometimes easier to use a proof by contradiction so that we can assume that the something exists.
%In symbolic form, we can say that we want to prove that
%\[
%\mynot \left( \exists x \in U \right) \left( P(x) \right),
%\]
%where $P(x)$ is some open sentence.  Instead of working with equivalent statement 
%$\left( \forall x \in U \right) \left( \mynot P(x) \right)$, it is sometimes easier to use a proof by contradiction and assume that $\left( \exists x \in U \right) \left( P(x) \right)$.  
For example, suppose we want to prove the following proposition:

\begin{proposition} \label{prop:persquare}
For all integers $x$ and $y$, if $x$ and $y$ are odd integers, then there does not exist an integer $z$ such that $x^2 + y^2 = z^2$.
\end{proposition}
Notice that the conclusion involves trying to prove that an integer with a certain property does not exist.  If we use a proof by contradiction, we can assume that such an integer $z$ exists.  This gives us more with which to work.
\begin{prog} \label{prog:persquare}
Complete the following proof of Proposition~\ref{prop:persquare}:
\begin{myproof}
We will use a proof by contradiction.  So we assume that there exist integers $x$ and $y$ such that $x$ and $y$ are odd and there exists an integer $z$ such that $x^2 + y^2 = z^2$.  Since $x$ and $y$ are odd, there exist integers $m$ and $n$ such that $x = 2m + 1$ and $y = 2n + 1$.
\begin{enumerate}
  \item Use the assumptions that $x$ and $y$ are odd to prove that $x^2 + y^2$ is even and hence, $z^2$ is even.  (See Theorem~\ref{T:n2odd} on page~\pageref{T:n2odd}.)
\end{enumerate}
We can now conclude that $z$ is even.  (See Theorem~\ref{T:n2odd} on page~\pageref{T:n2odd}.)  So there exists an integer $k$ such that $z = 2k$.  If we substitute for $x$, $y$, and $z$ in the equation $x^2 + y^2 = z^2$, we obtain
\[
(2m + 1)^2 + (2n + 1)^2 = (2k)^2.
\]
\setcounter{oldenumi}{\theenumi}
\begin{enumerate} \setcounter{enumi}{\theoldenumi}
  \item Use the previous equation to obtain a contradiction.  \hint One way is to use algebra to obtain an equation where the left side is an odd integer and the right side is an even integer. \qedhere
\end{enumerate}
\end{myproof}

\end{prog}
\hbreak

\endinput

