\subsection*{Cartesian Products}

When working with Cartesian products, it is important to remember that the Cartesian product of two sets is itself a set.  As a set, it consists of a collection of elements.  In this case, the elements of a Cartesian product are ordered pairs.  We should think of an ordered pair as a single object that consists of two other objects in a specified order.  For example,
\begin{itemize}
\item If  $a \ne 1$, then the ordered pair  $\left( {1,a} \right)$ is not equal to the ordered pair  $\left( {a,1} \right)$.  That is,  $\left( {1,a} \right) \ne \left( {a,1} \right)$.

\item If  $A = \left\{ {1,2,3} \right\}$  and  $B = \left\{ {a,b} \right\}$, then the ordered pair  $\left( {3,a} \right)$ is an element of the set   $A \times B$.  That is,  $\left( {3,a} \right) \in A \times B$.

\item If  $A = \left\{ {1,2,3} \right\}$  and  $B = \left\{ {a,b} \right\}$, then the ordered pair  $\left( {5,a} \right)$ is not an element of the set   $A \times B$ since $5 \notin A$.  That is,  $\left( {5,a} \right) \notin A \times B$.
\end{itemize}

In Section~\ref{S:setproperties}, we studied certain properties of set union, set intersection, and set complements, which we called the algebra of sets.  We will now begin something similar for Cartesian products.  We begin by examining some specific examples in Progress Check~\ref{prog:relationscartesian} and a little later in Progress Check~\ref{prog:cartprodintervals}.
\hbreak

\begin{prog}[\textbf{Relationships between Cartesian Products}] \hfill \label{prog:relationscartesian} \\
Let  $A = \left\{ {1,2,3} \right\}$, $T = \left\{ {1,2} \right\}$, $B = \left\{ {a,b} \right\}$
, and  $C = \left\{ {a,c} \right\}$.  We can then form new sets from all of the set operations we have studied.  For example,  $B \cap C = \left\{ a \right\}$, and so
\[
A \times \left( {B \cap C} \right) = \left\{ {\left( {1,a} \right),\left( {2,a} \right),\left( {3,a} \right)} \right\}.
\]

\begin{enumerate}
\item Use the roster method to list all of the elements (ordered pairs) in each of the following sets:
\label{A:relationscartesian1}%

\begin{multicols}{2}
\begin{enumerate}
  \item $A \times B$
  \item $T \times B$
  \item $A \times C$
  \item $A \times \left( {B \cap C} \right)$
  \item $\left( {A \times B} \right) \cap \left( {A \times C} \right)$
  \item $A \times \left( {B \cup C} \right)$
  \item $\left( {A \times B} \right) \cup \left( {A \times C} \right)$
  \item $A \times \left( {B - C} \right)$
  \item $\left( {A \times B} \right) - \left( {A \times C} \right)$
  \item $B \times A$
\end{enumerate}
\end{multicols}

\item List all the relationships between the sets in Part~(\ref{A:relationscartesian1}) that you observe.
\end{enumerate}
\end{prog}
\hbreak
