\section*{Section~\ref{S:recursion}}

\subsection*{Progress Check~\ref{prog:thirdfibonacci}}
\setcounter{equation}{0}
\begin{myproof}
We will use a proof by induction.  For each  natural number  $n$,  we let  $P( n )$
 be,
\begin{center}
$f_{3n} $  is an even natural number.
\end{center}

\noindent
Since  $f_3  = 2$, we see that  $P( 1 )$  is true and this proves the basis step.
\vskip6pt
\noindent
For the inductive step, we let  $k$  be a natural number and assume that  $P( k )$
 is true.  That is, assume that
$f_{3k} $  is an even natural number.
This means that there exists an integer  $m$  such that 
\begin{equation} \label{eq:progfiba}
f_{3k}  = 2m.
\end{equation}
We need to prove that  $P( {k + 1} )$ is true or that  $f_{3\left( {k + 1} \right)} $
 is even.  Notice that  $3( {k + 1} ) = 3k + 3$ and, hence,  
$f_{3\left( {k + 1} \right)}  = f_{3k + 3} $.  We can now use the recursion formula for the Fibonacci numbers to conclude that  
\[
f_{3k + 3}  = f_{3k + 2}  + f_{3k + 1} .
\]
Using the recursion formula again, we get $f_{3k + 2}  = f_{3k + 1}  + f_{3k}$.
Putting this all together, we see that
\begin{align}
  f_{3\left( {k + 1} \right)}  &= f_{3k + 3}  \notag \\ 
                               &= f_{3k + 2}  + f_{3k + 1}  \notag \\ 
                               &= \left( {f_{3k + 1}  + f_{3k} } \right) + f_{3k + 1}   \notag \\ 
                               &= 2f_{3k + 1}  + f_{3k}.  \label{eq:progfibb}
\end{align} 
We now substitute the expression for  $f_{3k} $ in equation~(\ref{eq:progfiba}) into 
equation~(\ref{eq:progfibb}).  This gives
\[
\begin{aligned}
  f_{3\left( {k + 1} \right)}  &= 2f_{3k + 1}  + 2m \\ 
  f_{3\left( {k + 1} \right)}  &= 2\left( {f_{3k + 1}  + m} \right) \\ 
\end{aligned} 
\]
This preceding equation shows that  $f_{3\left( {k + 1} \right)} $ is even.  Hence it has been proved that if  $P\left( k \right)$ is true, then $P\left( {k + 1} \right)$ is true and the inductive step has been established.  By the Principle of Mathematical Induction, this proves that  for each natural number  $n$,  the Fibonacci number  $f_{3n} $ is an even natural number.
\end{myproof}
\hbreak


\endinput
