\section*{Section~\ref{S:predicates} Open Sentences and Sets}
Plan about one class period for this section.

\subsection*{Main Topics}
Basic set notation including set-builder notation, variables and open sentences, the truth set of an open sentence, and set builder notation.

\subsection*{The Preview Activities}
\subsubsection*{Preview Activity~\ref{PA:sets} (Sets and Set Notation)} 
The purpose of this preview activity is simply to introduce students to an intuitive idea of how mathematicians use sets and the use of the roster method to designate the elements of a set.

\subsubsection*{Preview Activity~\ref{PA:variable} (Variables)}  
The concept of a \textbf{universal set} is introduced.  The purpose of problems two through seven is to provide a ``lead-in'' to the concept of a \textbf{truth set} introduced in the section.  This is also done in Progress Check~\ref{pr:predicates}.
\hbreak

%\subsubsection*{Activity~\ref{A:predicates}} This comes before the definition of a truth set. If students have read this section before class or if Preview Activity~\ref{PA:variables} seems to provide a sufficient introduction, then this activity can be skipped.

%\subsubsection*{Activity~\ref{A:truthset}}  This comes right after the definition of truth set and provides practice with working with this definition.  If students do not work on this activity, then it should be used to provide examples in class.  (Or the instructor can use other examples.
%\hbreak
%
\subsection*{The Exercises}

It is a good idea to assign most of the exercises in this section.  Exercise~(\ref{exer:sec21-3}) and~(\ref{exer:sec23-sets}) are needed to provide practice in using set-builder notation.  %Exercise~(\ref{exer:sec21-4}) is a good exercise since it makes the students distinguish between statements and open sentences.  Exercise~(\ref{Exer:quantifier}) is also good for making the distinction between statements and predicates.

\vskip6pt
\noindent
Typical Assignment:  Exercises 1(a, b, d, e), 2, 3, 4, 5, 6
\hbreak


\subsection*{Explorations and Activities}
This activity is intended to give students a better understanding of closure for a set with respect to an operation.  It is provides an opportunity to work with a universally quantified conditional statement and counterexamples for such statements.
\hbreak


\endinput
