\section{Inverse Functions} \label{S:inversefunctions}
\setcounter{previewactivity}{0}
%
For this section, we will use the concept of Cartesian product of two sets $A$ and $B$, denoted by $A \times B$, which is the set of all ordered pairs $(x, y)$ where $x \in A$ and $y \in B$.  That is,
\begin{center}
$A \times B = \left\{ { {\left( {x,y} \right)} \mid x \in A\text{ and }y \in B} \right\}.$
\end{center}
See \typeu Activity~\ref*{PA:cartesianproduct} in Section~\ref{S:cartesian} for a more thorough discussion of this concept. 
\input{previews/prev65-orderedpairs}
\input{previews/prev65-composition}
%
\setcounter{equation}{0}
\input{subsections/section65/sub65-orderedpair}
\input{subsections/section65/sub65-inversefunction}
\input{subsections/section65/sub65-notation}
\input{subsections/section65/sub65-theorems}
%\input{subsections/section65/sub65-constructinverse}

\endinput

%
%
%
\begin{activity}[Construction and Existence of Inverse Functions]\label{A:constructinverse}\hfill \\
\index{inverse of a function!construction}%
If $f\x A \to B$ is a bijection, then we know that its inverse is a function.  If  we are given a formula for the function  $f$, it may be desirable to determine a formula for the function  $f^{ - 1} $.  This can sometimes be done, while at other times it is very difficult or even impossible.
\vskip10pt
\noindent
\textbf{Construction of an Inverse Function} \\
Let  $f\x \R \to \R$ be defined by  $f( x ) = 2x^3  - 7$.

\begin{enumerate}
\item Sketch a graph of this function and use the graph to explain why the function is a bijection (one-to-one and onto).  A formal proof is not required.
\end{enumerate}

One way to attempt to find a formula for the inverse of this function is to set  
$y = f( x )$ and solve for  $x$.  In this case, we are using  $x$  for the input of  $f$  and  $y$  for the output of  $f$.  By solving for  $x$  in terms of  $y$, we are attempting to write a formula where   $y$  is the input and  $x$  is the output.  This formula represents the inverse function.

\begin{enumerate}
\setcounter{enumi}{1}
\item Solve the equation  $y = 2x^3  - 7$ for  $x$.  Use this to write a formula for  
$f^{ - 1} ( y )$, where  $f^{ - 1} \x \R \to \R$.  \label{A:constructinverse2}

\item Use the result of Part~(\ref{A:constructinverse2}) to verify that for each  
$x \in \R$,  $f^{ - 1} \!\left( {f( x )} \right) = x$ and for each  
$y \in \R$,  $f \!\left( {f^{ - 1} ( y )} \right) = y$.

\end{enumerate}

\noindent
\note  If we write  $y = f( x )$ for the function  $f$, we are using  $x$  as the independent variable (input) and  $y$  as the dependent variable (output).  In the case where  $f^{ - 1} $ is a function, we may write  $x = f^{ - 1} ( y )$, where  $y$  is the independent variable for  
$f^{ - 1} $.  However, when we are using real functions, we frequently want to graph them.  So we traditionally write each function as a function of  $x$.  
%Even if we have 
%\[
%f^{ - 1} ( y ) = \sqrt[3]{{\frac{{y + 7}}{2}}}, 
%\]
%we will frequently write 
%\[
%f^{ - 1} ( x ) = \sqrt[3]{{\frac{{x + 7}}{2}}}.
%\]
\vskip6pt
\noindent
\textbf{Existence of an Inverse Function}\index{inverse of a function!existence} \\
In calculus we learned that a function  $G$  is an  \textbf{increasing function}
\index{increasing function}%
\index{function!increasing}%
 provided that  
$G( {x_2 } ) > G( {x_1 } )$  whenever  $x_2  > x_1 $.  We also learned that a differentiable function  $G$  is increasing on an interval provided that  
$G'( x ) > 0$ for all  $x$  in that interval.

\begin{enumerate} \setcounter{enumi}{3}
\item Explain why an increasing function is a one-to-one function or an injection.
\end{enumerate}

\noindent
Let  $g\x \R \to \R$ be defined by  $g( x ) = x^3  + x - 5$.

\enlargethispage{\baselineskip}
\begin{enumerate}
\setcounter{enumi}{4}
\item Use  $g'( x )$ to show that $g$ is an injection, and use a graph of  $g$  to show that  $g$  is a surjection.  Explain why this proves the existence of an inverse function  
$g^{ - 1} \x \R \to \R$.

\item If  $y = g( x )$ and hence $y = x^3  + x - 5$, do you think it is possible to solve this equation for  $x$  to obtain a formula for  $g^{ - 1} ( y )$?  Explain.
\end{enumerate}
\end{activity}
\hbreak
%
\begin{activity}[The Inverse Sine Function] \label{A:inversesine} \hfill \\
In order to obtain an inverse function, it is sometimes necessary to restrict the domain (or the codomain) of a function.

\begin{enumerate}
\item Let  $f\x \R \to \R$ be defined by  $f( x ) = \sin x$.  Explain why the inverse of the function  $f$  is not a function.  (A graph may be helpful.)  \label{A:inversesine1}
\end{enumerate}

Notice that if we use the ordered pair representation, then the sine function can be represented as
\[
f  = \left\{ { {( {x, y} ) \in \R \times \R } \mid y = \sin x} \right\}\!.
\]
If we denote the inverse of the sine function by  $\sin ^{ - 1} $, then
\[
f^{ - 1}  = \left\{ { {( {y, x} ) \in \R \times \R } \mid y = \sin x} \right\}\!.
\]
Part~(\ref{A:inversesine1}) proves that  $f^{ - 1} $  is not a function.  However, in previous mathematics courses, we frequently used the ``inverse sine function.''  This is not really the inverse of the sine function as defined in Part~(\ref{A:inversesine1}) but, rather, it is the inverse of the \linebreak
\vspace{2pt}
\noindent
sine function \textbf{restricted to the domain}  
$\left[ {-\dfrac{{\pi }}{2}, \dfrac{\pi }{2}} \right]$.

\begin{enumerate}
\setcounter{enumi}{1}
\item Explain why the function  
$F\x \left[ {- \dfrac{{\pi }}{2}, \dfrac{\pi }{2}} \right] \to \left[ { - 1, 1} \right]$
defined by 
$F( x ) = \sin x$ is a bijection.  \label{A:inversesine2}
\end{enumerate}

The inverse of the function in Part~(\ref{A:inversesine2}) is itself a function and is called the \textbf{inverse sine function}
\index{inverse sine function}%
 (or sometimes the \textbf{arcsine function}).
\index{arcsine function}%

\begin{enumerate}
\setcounter{enumi}{2}
\item What is the domain of the inverse sine function?  What are the range and codomain of the inverse sine function?
\end{enumerate}
%
Let us now use  $F ( x ) = \text{Sin} ( x )$ 
\label{sym:restrictsine} to represent the restricted sine function in Part~(\ref{A:inversesine2}).  Therefore, 
$F^{-1} ( x ) = \text{Sin}^{ - 1} ( x ) $
\label{sym:inversesine}  can be used to represent the inverse sine function.  Observe that
\[
F\x \left[ -{\frac{{\pi }}{2}, \frac{\pi }{2}} \right] \to \left[ { - 1, 1} \right]
	\text{ and }	
F^{ - 1} \x \left[ { - 1, 1} \right] \to \left[ -{\frac{{\pi }}{2}, \frac{\pi }
{2}} \right].
\]
%
\begin{enumerate}
\setcounter{enumi}{3}
\item Using this notation, explain why
\begin{list}{}
\item $\text{Sin}^{ - 1} y = x	\text{ if and only if }	
\left[ {y = \sin x\text{  and  } -\dfrac{{\pi }}{2} \leq x \leq \dfrac{\pi }
{2}} \right]$;

\item $\text{Sin} \!\left( \text{Sin}^{-1} ( y ) \right) = y$ for all 
$y \in \left[ -1, 1 \right]$; and
\item $\text{Sin}^{-1} \!\left( \text{Sin} ( x ) \right) = x$ for all 
$x \in \left[ - \dfrac{\pi}{2}, \dfrac{\pi}{2} \right]$.
\end{list}
\end{enumerate}
\hbreak


\end{activity}

\endinput

\begin{center}
\setlength{\unitlength}{0.3cm}
\begin{picture}(14,10)
\put(2,4){\oval(4,6)}
\put(10,4){\oval(4,6)}

\put(2,2){\circle*{.5}}
\put(2,6){\circle*{.5}}

\put(9.7,4){\circle*{.5}}
\put(9.7,2.5){\circle*{.5}}

\put(1,1.8){$t$}
\put(1,5.8){$s$}
\put(10.4,3.8){$y$}
\put(10.4,2.3){$z$}

\put(1.8,8){$A$}
\put(10,8){$B$}

\put(3,8.2){\vector(1,0){6.6}}
\put(6,8.8){$f$}

\put(2.2,2){\vector(4,1){6.8}}
\put(2.2,6){\vector(4,-1){6.8}}

\end{picture}



\begin{center}
\setlength{\unitlength}{0.5cm}
\begin{picture}(22,10)
\put(2,5){\oval(3,6)}
\put(10,5){\oval(3,6)}
\put(18,5){\oval(3,6)}

\put(1.5,5){$A$}
\put(9.5,5){$B$}
\put(17.5,5){$C$}

\put(4,6){\vector(1,0){4}}
\put(12,6){\vector(1,0){4}}
\put(8,4){\vector(-1,0){4}}
\put(16,4){\vector(-1,0){4}}

\put(6,6.5){$f$}
\put(14,6.5){$g$}
\put(6,3){$f^{ - 1}$}
\put(14,3){$g^{-1}$}

\put(3,8.5){\vector(1,0){14}}
\put(17,1.5){\vector(-1,0){14}}

\put(8.5,9){$g \circ f$}
\put(8.5,0.5){$( g \circ f )^{-1}$}

\end{picture}
\end{center}

\end{center}
