\section[Proving Set Relationships]{Proving Set Relationships} \label{S:provingset}
%\markboth{Chapter~\ref{C:settheory}. Set Theory}{\ref{S:provingset}. Proving Set Relationships}
\setcounter{previewactivity}{0}
%
\input{previews/prev52-twosets}
\input{previews/prev52-venn}

\input{subsections/section52/sub52-choose}
\input{subsections/section52/sub52-equality}
\input{subsections/section52/sub52-disjoint}

\endinput

\begin{itemize}
  \item Preview Activity 1.
  \item Combine Preview Activities 2 and 3.
  \item After Proposition 4.8, show that $S \ne T$ by showing $S \not \subseteq T$.
  \item Progress Check 4.12.
  \item Progress Check 4.9.
  \item For Proposition 4.11, use $A - \left( A \cap B^c \right) = A \cap B$.
  \item Possibly the current Prop. 4.11 as a Progress Check.  Or delay this to a theorem in the next section.
  \item Possible Activity or addition:  $A - ( B \cap C ) = (A - B) \cap (A - C)$.  The sets are not equal but the right side is a subset of the left side.
  \item Disjoint sets.  Change Proposition 4.14 to a biconditional (see Activity 4.17).  Do the converse as a Progress Check or an exercise.
  \item Using the Choose an Element Method in a Different Setting as an activity.
\end{itemize}


\subsection*{Using the Choose-an-Element Method in a Different Setting}
We have used the choose-an-element method
\index{choose-an-element method}%
 to prove Propositions~\ref{P:SissubsetT}, \ref{P:setdifference}, \linebreak
and~\ref{P:subsetprop}.  Proofs involving sets that use this method are sometimes referred to as \textbf{element-chasing proofs.}
\index{element-chasing proof}%
\index{proof!element-chasing}%
  This name is used since the basic method is to choose an arbitrary element from one set and ``chase it'' until you prove it must be in another set.

The choose-an-element method, however, is a general proof technique and can be used in settings other than set theory.  It is often used whenever we encounter a universal quantifier in a statement in the backward process.  Consider the following proposition.
\begin{proposition} \label{P:divlinearcomb}
Let a, b, and  t  be integers with $t \ne 0$.  If  t  divides  a  and  t  divides  b, then for all integers  x  and  y,  t  divides  (ax + by).
\end{proposition}
%
Notice that the conclusion of the conditional statement in this proposition involves the universal quantifier.  So in the backward process, we would have
\begin{list}{}
\item $Q$:	For all integers  $x$  and  $y$, $t$  divides  $ax + by$.
\end{list}
\vskip10pt
%
The ``elements'' in this sentence are the integers  $x$  and  $y$.  In this case, these integers have no ``given property.''  (They are just integers.)  The ``something that happens'' is that 
$t$  divides  $\left(ax + by\right)$.  
%Of course, this quantified sentence could be written as a conditional statement as follows:
%\begin{list}{}
%\item $Q$:	If  $x$  and  $y$  are integers, then  $t$  divides  $\left(ax + by\right)$.
%\end{list}
%\vskip10pt
%
This means that in the forward process, we can use the hypothesis of the proposition and choose integers  $x$  and  $y$.  That is, in the forward process, we could have
\begin{list}{}
\item $P$:	$a$, $b$, and  $t$  are integers with $t \ne 0$, $t$  divides  $a$  and  $t$  divides  $b$.
\item $P1$:	Let  $x \in \Z$ and let  $y \in \Z$.
\end{list}
\vskip10pt
%
\noindent
The proof of this proposition is part of Activity~\ref{A:divlincomb}.
\hbreak

\begin{activity}[Exploring and Proving Proposition~\ref{P:divlinearcomb}] \label{A:divlincomb} \hfill
\begin{enumerate}
\item Whenever we encounter a new proposition, it is a good idea to explore the proposition by looking at specific examples.  For example, let  \linebreak
$a = 20$, $b = 12$, and  $t = 4$.  In this case,  $t \mid a$  and  $t \mid b$.  In each of the following cases, determine the value of  $\left(ax + by\right)$ and determine if $t$ divides 
$\left(ax + by\right)$.  \label{A:divlincomb1}

\begin{multicols}{3}
\begin{enumerate}
\item $x=1, y=1$
\item $x=1, y=-1$
\item $x=2, y=2$
\item $x=2, y=-3$
\item $x=-2, y=3$
\item $x=-2, y=-5$
\end{enumerate}
\end{multicols}

\item Repeat Part~(\ref{A:divlincomb1}) with  $a = 21$, $b =  - 6$, and  $t = 3$.

\item We started the forward-backward process for the proof of Proposition~\ref{P:divlinearcomb} following the discussion of this proposition.  Complete the following proof of Proposition~\ref{P:divlinearcomb}.
%
\addtocounter{theorem}{-2}
\begin{proposition}
Let a, b, and  t  be integers with $t \ne 0$.  If  t  divides  a  and  t  divides  b, then for all integers  x  and  y,  t  divides  ax + by.
\end{proposition}
%
\begin{myproof}
Let $a$, $b$, and  $t$  be integers with $t \ne 0$, and assume that $t$  divides  $a$  and  $t$  divides  $b$.  We will prove that for all integers  $x$  and  $y$,  $t$  divides  
$\left(ax + by\right)$.

So let  $x \in \Z$ and let  $y \in \Z$.  Since  $t$  divides  $a$, there exists an integer  $m$  such that $ \ldots .$
\end{myproof}
\end{enumerate}
\end{activity}
\hbreak
\addtocounter{theorem}{1}

\begin{activity}[The Converse of Proposition~\ref{P:subsetprop}] \label{A:converseprop} \hfill \\
We will explore the converse of Proposition~\ref{P:subsetprop}.  This converse states:

\begin{list}{}
\item \emph{Let $A$ and $B$ be subsets of some universal set $U$\!.  If $A \cap B^c = \emptyset$, then $A \subseteq B$}.
\end{list}

\begin{enumerate}
\item Let $U = \Z$ and let $A = \left\{ x \in \Z \mid x \geq 8 \right\}$ and 
$T = \left\{ x \in \Z \mid x \leq 2 \right\}$.  We then see that $A \cap T = \emptyset$.  So if we let $B$ be the subset of $\Z$ such that $B^c = T$, we then have an example of two sets such that $A \cap B^c = \emptyset$.  In this case, is $A$ a subset of $B$?  Explain.

\item Give an example of two sets $A$ and $B$ that are subsets of $\R$ and for which 
$A \cap B^c = \emptyset$.  In this case, is $A$ a subset of $B$?  Explain.

\item Is the converse of Proposition~\ref{P:subsetprop} true or false?  Justify your conclusion.
\end{enumerate}
\end{activity}
\hbreak




\endinput
