\section{Constructing Direct Proofs} \label{S:direct}
%\markboth{Chapter \ref{C:intro}. Introduction to Writing}{\ref{S:direct}. Constructing Direct Proofs}
\setcounter{previewactivity}{0}
%\input{focus/focus12}
\input{previews/prev12-evenodd}
\input{previews/prev12-firstproof}


\input{subsections/section12/sub12-properties-numbers}
%


\input{subsections/section12/sub12-construct-proof}



\input{subsections/section12/sub12-writing}



\input{subsections/section12/sub12-comments}



\begin{prog}[\textbf{Exploring a Proposition}]\label{A:kstable2} \hfill \\
Construct a table of values for  $\left( {3m^2  + 4m + 6} \right)$
 using at least six different integers for  $m$.  Make one-half of the values for  $m$  even integers and the other half odd integers.  Is the following proposition true or false?  

\begin{center}
If $m$ is an odd integer, then $\left(3m^2 + 4m + 6 \right)$ is an odd integer.
\end{center}
Justify your conclusion.  This means that if the proposition is true, then you should write a proof of the proposition.  If the proposition is false, you need to provide an example of an odd integer for which $\left(3m^2 + 4m + 6 \right)$ is an even integer.
\end{prog}
\hbreak


\begin{prog}[\textbf{Constructing and Writing a Proof}] \label{pr:pythag} \hfill \\
The \textbf{Pythagorean Theorem}
\index{Pythagorean Theorem}%
for right triangles states that if $a$ and $b$ are the lengths of the legs of a right triangle and $c$ is the length of the hypotenuse, then $a^2 + b^2 = c^2$.  For example, if $a = 5$ and $b = 12$ are the lengths of the two sides of a right triangle and if $c$ is the length of the hypotenuse, then the $c^2 = 5^2 + 12^2$ and so 
$c^2 = 169$.  Since $c$ is a length and must be positive, we conclude that $c = 13$.

\newpar
Construct and provide a well-written proof for the following proposition.

\newpar
\textbf{Proposition}.  If $m$ is a real number and $m$, $m + 1$, and $m + 2$ are the lengths of the three sides of a right triangle, then $m = 3$.

\newpar
Although this proposition uses different mathematical concepts than the one used in this section, the process of constructing a proof for this proposition is the same forward-backward method that was used to construct a proof for Theorem~\ref{T:xyodd}.  However, the backward question, ``How do we prove that $m = 3$?'' is simple but may be difficult to answer.  The basic idea is to develop an equation from the forward process and show that $m = 3$ is a solution of that equation.
\end{prog}
\hbreak


\endinput




