\section*{Section~\ref{S:mathinduction} The Principle of Mathematical Induction}

\subsection*{Preview Activity 1 \textbf{(Exploring Statements of the Form \mathversion{bold} $\left( \forall n \in \N \right)\left( P(n) \right)$)}}
The following table shows that $P(1)$ through $P(7)$ are true.
\begin{center}
\begin{tabular}[t]{| c | c | c |} \hline
$n$  &  $\left( {5^n - 1} \right)$  &  Does 4 divide $\left( {5^n - 1} \right)$? \\ \hline
1  &  4  &  yes  \\ \hline
2  &  24  &  yes \\ \hline
3  &  124  &  yes  \\ \hline
4  &  624  &  yes  \\ \hline
5  &  3124  &  yes \\ \hline
6  &  15,624  &  yes  \\ \hline
7  &  78,124  &  yes  \\ \hline
\end{tabular}
\end{center}
%Based on the work summarized in this table, it seems the following is a true proposition:
%\begin{center}
%For each natural number  $n$, 4  divides $5^n - 1$.
%\end{center}
The following table shows that $Q(1)$ through $Q(7)$ are true.
\begin{center}
\begin{tabular}[t]{| c | c | c |} \hline
$n$  &  $1^2  + 2^2  + \, \cdots \, + n^2 $  &  $\dfrac{{n(n + 1)(2n + 1)}}{6}$ \\ \hline
1  &  1  &  1  \\ \hline
2  &  5  &  5 \\ \hline
3  &  14  &  14  \\ \hline
4  &  30  &  30  \\ \hline
5  &  55  &  55 \\ \hline
6  &  91  &  91  \\ \hline
7  &  140  &  140  \\ \hline
\end{tabular}
\end{center}

\hbreak

\subsection*{Preview Activity 2 \textbf{(A Property of the Natural Numbers)}}
\begin{enumerate}
  \item A set $T$ that is a subset of $\Z$ is not an inductive set provided that there exists an integer $k$ such that $k \in T$ and $k+1 \notin T$
\item \begin{enumerate}
\item The set  $A = \left\{ {1, 2, 3,  \ldots , 20} \right\}$  is not inductive since  
$20 \in A$, but $21 \notin A$.

\item The set of natural numbers, $\mathbb{N}$, is inductive since the natural numbers are closed with respect to addition.


\item The set  $B = \left\{ {\left. {n \in \mathbb{N}} \right| n \geq 5} \right\}$
 is inductive.  If  $k \in \mathbb{Z}$ and  $k \geq 5$, then  
$\left( {k + 1} \right) \geq 5$.

\item The set  $S = \left\{ {\left. {n \in \mathbb{N}} \right| n \geq  - 3} \right\}$
 is inductive.  If  $k \in \mathbb{Z}$ and  $k \geq  - 3$, then  
$\left( {k + 1} \right) \geq  - 3$.

\item The set  $R = \left\{ {n \in \mathbb{Z}\left| {n \leq 100} \right.} \right\}$
 is not inductive since $100 \in R$ but $101 \notin R$.

\item The set of integers, $\mathbb{Z}$, is inductive since  the set of integers is closed with respect to addition.
\item The set of all odd integers is not an inductive set.  For example, $1$ is in the set of all odd integers but 2 is not in the set of all odd integers.
\end{enumerate}


\item Now assume that  $T \subseteq \mathbb{N}$ and assume that  $1 \in T$ and that  $T$ is inductive.
\begin{enumerate}
\item Since $1 \in T$ and $T$ is inductive,   $2 \in T$.

\item Since $2 \in T$ and $T$ is inductive,  $3 \in T$.

\item Since $3 \in T$ and $T$ is inductive,   $4 \in T$.

\item Since $4 \in T$ and $T$ is inductive, $5 \in T$.  This in turn guarantees that  $6 \in T$, which guarantees that  $7 \in T$.  We can continue this process to argue that any succeeding natural number is in  $T$.  So, $100 \in T$.

\item The argument in Part (d) suggests that  $T = \mathbb{N}$.

\end{enumerate}
\end{enumerate}
\hbreak



\newpage

\endinput
