\section*{Section~\ref{S:quantifier} Quantifiers and Negations}
\subsection*{Preview Activity 1 (An Introduction to Quantifiers)}
\begin{enumerate}
    \item $\left( \forall a \in \mathbb{R}\right) \left(a + 0 = a\right)$.  This is a true statement.
    \item $3x - 5 = 9$.  This is an open sentence whose truth set is $\left\{ \dfrac{14}{3} \right\}$.
    \item $\sqrt x  \in \mathbb{R}$.  This is an open sentence whose truth set is the set of nonnegative real numbers.
    %\item $\left( \forall x \in \mathbb{R}\right) \left( \sin( {2x}) = 2( {\sin x})( {\cos x})$\right).
    \item $\sin( {2x} ) = 2( {\sin x} )( {\cos x})$.  This is an open sentence whose truth set is the set of real numbers.
    \item $\left( \forall x \in \R\right) \left(\sin( {2x} ) = 2( {\sin x})( {\cos x}) \right)$.  This is a true statement.
    \item $\left( \exists x \in \R \right)\left( x^2  + 1 = 0 \right)$.  This is a false statement.
    \item $\left( \forall x \in \R \right) \left( x^3  \geq x^2 \right)$.  This is a false statement.  A counterexample is $x = 0.5$.
    \item $x^2  + 1 = 0$.  This is an open sentence.  If the universal set is $\R$, then the truth set is the empty set.
    \item If  $x^2 \geq 1$, then  $x  \geq 1$.  This is an open sentence whose truth set is 
$\left\{ x \in \R \mid x > -1 \right\}$.
    \item $\left( \forall x \in \Z \right)\left( \text{If } x^2 \geq 1, \text{ then } x \geq 1 \right)$.  This is a false statement.  A counterexample is $x = -1$.
\end{enumerate}
\hbreak



\subsection*{Preview Activity 2 (Attempting to Negate Quantified Statements)}
\begin{enumerate}
\item \begin{enumerate}
  \item Every integer is even.
  \item The statement is false since there are integers that are not even.
  \item There exists an integer that is not even.
  \item $( {\exists x \in \mathbb{Z}} )( {x\text{ is not even}} )$.
\end{enumerate}

\item \begin{enumerate}
  \item There exists a real number $x$ such that $x^3 > 0$.
  \item The statement is true since there are real numbers whose cube is positive.  For examaple, if $x = 1$, then $x^3 > 0$.
  \item For each real number $x$, $x^3 \leq 0$.
  \item $\left( \forall x \in \R \right) \left( x^3 \leq 0 \right)$.
  \end{enumerate}
\end{enumerate}
\hbreak


\newpage

\endinput
