\begin{previewactivity}[\textbf{Truth Values of Statements}]\label{PA:truth} \hfill \\
We will use the following two statements for all of this activity:
\begin{itemize}
\item $P$  is the statement ``It is raining.'' 
\item $Q$  is the statement ``Daisy is playing golf.''
\end{itemize}
In each of the following four parts, a truth value will be assigned to statements $P$ and 
$Q$.  For example, in Question~(1), we will assume that each statement is true.  In Question~(2), we will assume that $P$ is true and $Q$ is false.  In each part, determine the truth value of each of the following statements:
\renewcommand{\labelenumi}{(\textbf{\alph{enumi}})}
  \begin{enumerate}
    \item \makebox[1.2in][l]{($P \wedge Q$)} It is raining and Daisy is playing golf.
    \item \makebox[1.2in][l]{($P \vee Q$)}  It is raining or Daisy is playing golf.
    \item \makebox[1.2in][l]{($P \to Q$)}	If it is raining, then Daisy is playing golf.
    \item \makebox[1.2in][l]{($\mynot P$)}	It is not raining.
    %\item \makebox[1.2in][l]{(not $P$ or $Q$)}	It is not raining or Daisy is playing golf.
    %\item \makebox[1.2in][l]{($P$ and not $Q$)}	It is raining and Daisy is not playing golf.
  \end{enumerate}
\renewcommand{\labelenumi}{\textbf{\arabic{enumi}.}}

\newpar
Which of the four statements [(a) through (d)] are true and which are false in each of the following four situations?
\begin{enumerate}
\item When $P$ is true (it is raining) and  $Q$ is true (Daisy is playing golf).  

\item When  $P$ is true (it is raining) and  $Q$ is false (Daisy is not playing golf).   

\item When  $P$ is false (it is not raining) and  $Q$ is true (Daisy is playing golf).    

\item When  $P$ is false (it is not raining) and  $Q$ is false (Daisy is not playing golf).    
\end{enumerate}
\hbreak
\end{previewactivity}
