\begin{previewactivity}[Working with Venn Diagrams] \label{PA:workingwithvenn2} \hfill \\
Let $A$ and $B$ be subsets of some universal set $U$.

\begin{enumerate}
\item Draw two general Venn diagrams for the sets  $A$  and  $B$.  On one, shade the region that represents  $\left( {A \cup B} \right)^c $, and on the other, shade the region that represents  $A^c  \cap B^c $.  Explain carefully how you determined these regions. 
\label{PA:workingwithvenn2-1}%

\item Based on the Venn diagrams in Part~(\ref{PA:workingwithvenn2-1}), what appears to be the relationship between the sets   $\left( {A \cup B} \right)^c $ and   $A^c  \cap B^c $?

\item Draw two general Venn diagrams for the sets  $A$  and  $B$.  On one, shade the region that represents  $\left( {A \cap B} \right)^c $, and on the other, shade the region that represents  $A^c \cup B^c $.  Explain carefully how you determined these regions.  \label{PA:workingwithvenn2-3}

\item Based on the Venn diagrams in Part~(\ref{PA:workingwithvenn2-3}), what appears to be the relationship between the sets      $\left( {A \cap B} \right)^c $ and  $A^c \cup B^c $?

\end{enumerate}
\end{previewactivity}
\hbreak
%
\begin{previewactivity}[Working with Definitions] \label{PA:workingwithdef} \hfill \\
Some of the properties of set operations are closely related to some of the logical operators we studied in Section~\ref{S:logop}.  This is due to the fact that set intersection is defined using a conjunction (and), and set union is defined using a disjunction (or).  For example, if  $A$  and  $B$ are subsets of some universal set  $U$, then an element  $x$  is in  $A \cup B$  if and only if  $x \in A$  or  $x \in B$.


%consider one of De Morgan's Laws for statements  $P$  and  $Q$.
%\[
%\mynot \left( {P \vee Q} \right) \equiv \mynot P \wedge \mynot Q.
%\]
%
%Now, if  $A$  and  $B$ are subsets of some universal set  $U$, then an element  $x$  is in  $A \cup B$  if and only if  $x \in A$  or  $x \in B$.

\begin{enumerate}
  \item Use one of De Morgan's Laws (Theorem~\ref{T:logequiv} on page~\pageref{T:logequiv}) to explain carefully what it means to say that an element  $x$  is not in  $A \cup B$. 
\label{part1PA1}
  \item What does it mean to say that an element  $x$  is in  $A^c $? What does it mean to say that an element  $x$  is in  $B^c $? 
  \item Explain carefully what it means to say that an element  $x$   is in  $A^c  \cap B^c $. \label{part3PA1}
  \item Compare your response in Part~(\ref{part1PA1}) to your response in Part~(\ref{part3PA1}).  Are they equivalent?  Explain.
  \item How do you think the sets  $\left( {A \cup B} \right)^c $ and $A^c  \cap B^c $ are related?  Is this consistent with the Venn diagrams from Preview Activity~\ref{PA:workingwithvenn2}?
\end{enumerate}
\end{previewactivity}
\hbreak
%

\begin{previewactivity}[Proving that Statements Are Equivalent] \label{PA:provingequiv} \hfill

\begin{enumerate}
\item Let $X$, $Y$, and $Z$ be statements.  Complete a truth table for  \\$\left[ {\left( {X \to Y} \right) \wedge \left( {Y \to Z} \right)} \right] \to \left( {X \to Z} \right)$.

\item Assume that  $P$, $Q$, and  $R$  are statements and that we have proven that the following conditional statements are true:
\begin{multicols}{2}
\begin{itemize}
  \item If  $P$  then  $Q$.  $\left( {P \to Q} \right)$
  \item If  $Q$  then  $R$.  $\left( {Q \to R} \right)$
  \item If  $R$  then  $P$.  $\left( {R \to P} \right)$
\end{itemize}
\end{multicols}
Explain why each of the following statements is true.
\begin{multicols}{2}
\begin{enumerate}
  \item $P$  if and only if  $Q$.  $\left( {P \leftrightarrow Q} \right)$.
  \item $Q$  if and only if  $R$.	$\left( {Q \leftrightarrow R} \right)$.
  \item $R$  if and only if  $P$.  $\left( {R \leftrightarrow P} \right)$.
\end{enumerate}
\end{multicols}
Remember that  $X \leftrightarrow Y$  is logically equivalent to  
$\left( {X \to Y} \right) \wedge \left( {Y \to X} \right)$.

\end{enumerate}
\end{previewactivity}
\hbreak


\endinput
