\begin{previewactivity}[\textbf{An Equation with Two Variables}] \label{PA:equationwith2vars} \hfill \\
In Section~\ref{S:predicates}, we introduced the concept of the \textbf{truth set of an open sentence with one variable.}
\index{truth set}%
  This was defined to be the set of all elements in the universal set that can be substituted for the variable to make the open sentence a true statement.

In previous mathematics courses, we have also had experience with open sentences with two variables.  For example, if we assume that  $x$  and  $y$  represent real numbers, then the equation 
\[
2x + 3y = 12
\]
is an open sentence with two variables.  An element of the truth set of this open sentence (also called a solution of the equation) is an ordered pair  $\left( {a,b} \right)$ of real numbers so that when  $a$  is substituted for  $x$  and  $b$  is substituted for  $y$, the open sentence becomes a true statement (a true equation in this case).  For example, we see that the ordered pair $(6, 0)$ is in the truth set for this open sentence since
\[
2 \cdot 6 + 3 \cdot 0 = 12
\]
is a true statement.  On the other hand, the ordered pair $(4, 1)$ is not in the truth set for this open sentence since
\[
2 \cdot 4 + 3 \cdot 1 = 12
\]
is a false statement.

\noindent
\textbf{Important Note}: The order of the  two numbers in the ordered pair is very important.  We are using the convention that the first number is to be substituted for $x$ and the second number is to be substituted for $y$.  With this convention, $\left( 3, 2 \right)$ is a solution of the equation $2x + 3y = 12$, but $\left( 2, 3 \right)$ is not a solution of this equation.

\begin{enumerate}
\item List six different elements of the truth set (often called the solution set) of the open sentence 
 with two variables $2x + 3y = 12$.

\item From previous mathematics courses, we know that the graph of the equation $2x + 3y = 12$ is a straight line.  Sketch the graph of the equation  $2x + 3y = 12$ in the $xy$-coordinate plane.  What does the graph of the equation  $2x + 3y = 12$ show?

\item Write a description of the solution set  $S$  of the equation  $2x + 3y = 12$ using set builder notation.

\end{enumerate}
\end{previewactivity}
\hbreak

\endinput
